factor_explorer

### 2014-06-20 ijlyttle/factor_explorer.Rmd
a possibly-ill-fated experiment to make a data-exploration tool

# First pass through the data

This will be a stream-of-concoiusness walk through the data, a first pass to see what I can see.

@@

```{r load_libraries, message=FALSE}
library(stringr)
library(ggplot2)
library(dplyr)
library(tidyr)
library(lubridate)
library(ggvis)
```

@@

I suspect that for publication, `ggplot2` will be more capable than `ggvis`.

# Load the data

@@

Looking at the number of variables in each of the data frames, I wish I knew how to use ggobi...

One of the tracks to investigate is inequality. I see there is a package, `ineq` on CRAN, that claims to have a Gini-coefficient function, so that seems like a good place to start.

```{r ineq}
library(ineq)
```

Let's get an idea, for each country and subject, an estimate of the inequality of outcomes.

NOTE: `dplyr` seems to choke on grouping of data.frames with NA, hence the `filter()`.

Also: is there a problem the help text for `dplyr::grouped_df`? The description of `drop` seems backwards to me.

```{r twiddle}
df_twiddle <- student2012 %>%
  select(country = CNT, age = ST06Q01,
         math = PV1MATH, reading = PV1READ, science = PV1SCIE) %>%
  gather(key = subject, value = score, -country, -age) %>%
  do(filter(., complete.cases(.))) %>%
  dplyr::group_by(country, subject) %>%
  summarize(
    n = n(), 
    gini_coeff = ineq(score, type="Gini"), 
    mean = mean(score), 
    median = median(score)) %>%
  arrange(as.character(country))
```

"Alabama first" works to look for a specific country, but lets arrange by something more meaningful.

Let's get a mean of coefficents over all subjects, to help us order things.

```{r by_country}
df_twiddle_mean <- df_twiddle %>%
  group_by(country) %>%
  summarize(
    gini_coeff = mean(gini_coeff),
    mean = mean(mean),
    median = mean(median)
  )
```

```{r arrange}
my_levels <- 
  df_twiddle_mean$country[rev(order(df_twiddle_mean$gini_coeff))] %>%
  as.character

df_twiddle$country <- factor(
  df_twiddle$country,
  levels = my_levels
)

head(my_levels)

tail(my_levels)
```

I know I should figure out the `ggmap` package (I'll look at Di's code), but in the meantinme:

```{r vis_ineq_median}
ggplot(
  df_twiddle %>% filter(subject != "mean_subject"), 
  aes(x = gini_coeff, y = median, color = subject, size = n)
) + 
geom_point(alpha = 0.75) + 
scale_size_area()
```

This is print, so using ggvis and toolips is out of the question...

There appear two clusters - it will be interesting to see if there is a thread to bind them.

```{r vis_countries}
ggplot(
  df_twiddle,
  aes(x = gini_coeff, y = country, color = subject, size = n)
) + 
geom_point() + 
scale_size_area()

ggplot(
  df_twiddle,
  aes(x = median, y = country, color = subject, size = n)
) + 
geom_point() + 
scale_size_area()
```

Hopefully enough to get started.

>>>>

---
title: "factor_explorer"
runtime: shiny
output: html_document
---

## Intro

ggvis is amlmost there, but not quite, so I'll use ggplot.

@@

```{r load_libraries, message=FALSE}
library(stringr)
library(ggplot2)
library(dplyr)
library(tidyr)
library(lubridate)
library(shiny)
```

@@

@@

## Shiny app

We are told that we can put a shiny app into an Rmarkdown document - let's try that.

The thought is that we could use this (I wish we had another couple weeks) to discover the interesting relationships, then turn those into posters.

```{r shiny_variable_lists}
var_names_student <- dict_student2012$variable
names(var_names_student) <- 
  dict_student2012$description %>%
  str_replace_all("\x92", "'") # gets rid of funny characters"
```

```{r shiny_inputs}
selectInput(
  "test",
  "variable",
  choices = var_names_student
)
```
